---
title: "`r emo::ji('light_rail')` Commuters in Germany"
output: 
  flexdashboard::flex_dashboard:
    orientation: columns
    vertical_layout: fill
    theme: bootstrap
    css: "css/fonts.css"
    navbar:
      - { title: "<span class='fa fa-github' style='margin-right: -15px;'></span>",
          href: "https://github.com/CorrelAid/correlaidx-challenge-bremen",
          align: right }
    source_code: embed
runtime: shiny_prerendered
---

```{r include=FALSE, eval=FALSE}
# Python script to collect data using the Datenguide API:
# https://github.com/CorrelAid/correlaidx-challenge-bremen/blob/master/scripts/get_data.py
#
# This is only an interim solution. Ideally, the data would be queried directly
# when running the dashboard. That is technically doable with reticulate.
# 
# However, using datenguidepy to query data inside hosted shiny apps/documents is
# not possible at the moment because the shinyapps.io server does not have
# Python>=3.6.1, which is a requirement for pandas>=1.0.0 and datenguidepy.
```

<!-- Import stylesheets -->
<link href="https://code.cdn.mozilla.net/fonts/fira.css" rel="stylesheet"> 
<link href="https://code.cdn.mozilla.net/fonts/zilla-slab.css" rel="stylesheet">
<link href="https://cdn.jsdelivr.net/gh/tonsky/FiraCode@5/distr/fira_code.css" rel="stylesheet">

```{r setup, include=FALSE}
library(flexdashboard)
library(dplyr)
library(reactable)
library(htmltools)
library(highcharter)
```

```{r data, include=FALSE}
nuts1_geojson <- download_map_data("countries/de/de-all")
nuts3_geojson <- download_map_data("countries/de/de-all-all")
nuts1_map_data <- get_data_from_map(nuts1_geojson)

laender_df <- data.table::fread("laender_df.csv", encoding = "UTF-8",
                                colClasses = c(id = "character")) %>% 
  as_tibble() %>% 
  mutate(across(where(is.numeric), as.integer))

kreise_df <- data.table::fread("kreise_df.csv", encoding = "UTF-8",
                               colClasses = c(id = "character")) %>% 
  as_tibble() %>% 
  tidyr::drop_na() %>% # Osterode am Harz
  mutate(across(where(is.numeric), as.integer)) %>% 
  # Create key variable to join with highcharts map data
  mutate(parent_id = substr(id, 1, 2)) %>% 
  left_join(laender_df %>% select(id, bundesland = name) %>% distinct(),
            by = c("parent_id" = "id")) %>% 
  left_join(nuts1_map_data %>% select(name, `hc-key`),
            by = c("bundesland" = "name")) %>% 
  mutate(`hc-key` = paste0(`hc-key`, "-", id, "000"),
         # Fix unmatched keys
         `hc-key` = gsub("de-hb-04011000", "de-ni-04011000", `hc-key`),
         `hc-key` = gsub("de-ni-03159000", "de-ni-03152000", `hc-key`),
         `hc-key` = gsub("de-sn-14730000", "de-st-14730000", `hc-key`))
```

Sidebar {.sidebar}
=======================================================================

Data from the employment statistics of the [German Federal Employment Agency](https://www.arbeitsagentur.de/). Access to official data provided by the [datenguide API](https://github.com/datenguide/datenguide-api) via the [datenguidepy](https://github.com/CorrelAid/datenguide-python) package

---

<br>

```{r ui-input}
sliderInput("year", label = "Year:",
            min = 2011, max = 2019, value = 2019, step = 1, ticks = FALSE, sep = "")

selectInput("plot_var", label = "Statistic to plot:",
            choices = c("In", "Out", "Balance"), selected = "Balance")

radioButtons("choose_pct", label = "As a percentage of the population?",
             choices = c("Yes", "No"), selected = "No")
```

<br>

---

```{r download-ui}
selectInput("dataset", "Download data as CSV:",
            choices = c("NUTS-1", "NUTS-3"))

# downloadButton does not show up in Rmd documents
# https://stackoverflow.com/a/55725085
downloadButtonRmd <- function (outputId, label = "Download", class = NULL, ...) {
  tags$a(id = outputId,
         class = paste("btn btn-default shiny-download-link", class),
         href = "", target = "_blank", download = NA, 
         icon("download"), label, ...)
}

downloadButtonRmd("download_data", "Download")
```

```{r download-srvr, context="server"}
dataset_input <- reactive({
  switch(input$dataset,
         "NUTS-1" = laender_df,
         "NUTS-3" = kreise_df)
})

output$download_data <- downloadHandler(
  filename = function() {
    paste0("commuters_", input$dataset, ".csv")
  },
  content = function(file) {
    data.table::fwrite(dataset_input(), file)
  }
)
```

```{r prep-df, context="server"}
transform_stats <- function(data, to_pct = input$choose_pct) {
  if (to_pct == "Yes") {
    data %>% 
      mutate(across(c(In, Out, Balance), ~ round(.x * 100 / pop, 2)))
  } else {
    data
  }
}

laender_df_trans <- reactive({
  laender_df %>% 
    transform_stats()
})

kreise_df_trans <- reactive({
  kreise_df %>% 
    transform_stats()
})

# "Helper" to add percent sign to values
number_unit <- reactive({
  if (input$choose_pct == "No") ""
  else "%"
})
```

Between Bundesländer (NUTS-1) {data-icon="fa-th-list"}
=======================================================================

Column {data-width=7, .tabset}
-----------------------------------------------------------------------

### `r emo::ji("clipboard")` Table

```{r barchart-tbl-fn, context="server"}
# https://glin.github.io/reactable/articles/cookbook/cookbook.html#bar-charts-1
bar_chart <- function(label, width = "100%", height = "14px", fill = "#cccccc",
                      align = "left") {
  label <- div(style = list(marginLeft = "5px", marginRight = "5px"), label)
  bar <- div(style = list(background = fill, width = width, height = height,
                          float = align))
  if (align == "right") {
    div(style = list(display = "flex", alignItems = "center",
                     justifyContent = "flex-end"), label, bar)
  } else {
    div(style = list(display = "flex", alignItems = "center"), bar, label)
  }
}
```

```{r laender-tbl-ui}
reactableOutput("laender_tbl")
```

```{r laender-tbl-srvr, context="server"}
output$laender_tbl <- renderReactable({
  laender_df_trans() %>% 
    filter(year == input$year) %>% 
    select(name, Out, Balance, In) %>% 
    reactable(
      defaultPageSize = 16, borderless = TRUE, wrap = FALSE,
      defaultSorted = "Balance", defaultSortOrder = "desc",
      columns = list(
        name = colDef(name = "Bundesland", defaultSortOrder = "asc",
                      maxWidth = 250),
        Out = colDef(
          cell = function(value) {
            width <- paste0(value * .75 * 100 / max(.$In), "%")
            value <- paste0(format(value, big.mark = " "), number_unit())
            bar_chart(value, width = width, fill = "#c7522b", align = "right")
          },
          align = "right"
        ),
        In = colDef(
          cell = function(value) {
            width <- paste0(value * .75 * 100 / max(.$In), "%")
            value <- paste0(format(value, big.mark = " "), number_unit())
            bar_chart(value, width = width, fill = "#008585", align = "left")
          },
          align = "left"
        ),
        Balance = colDef(
          name = emo::ji("balance_scale"),
          width = 80,
          cell = function(value) paste0(format(value, big.mark = " "), number_unit()),
          style = list(background = "#f8f8f8"),
          align = "right"
        )
      )
    )
})
```

### `r emo::ji("world_map")` Choropleth map

```{r mapcols-fn, context="server"}
hc_map_colors <- function(hc_map, var, pct, level) {
  ret <- hc_map %>% 
      hc_colorAxis(labels = list(overflow = "alow"))
  if (var == "Balance") {
    ret <- ret %>% 
      hc_colorAxis(stops = color_stops(
        colors = colorspace::divergingx_hcl(10, rev = TRUE)
      ))
    if (pct == "No") {
      if (level == "nuts1") {
        ret <- ret %>% 
          hc_colorAxis(min = -2e5, max = 2e5)
      }
      if (level == "nuts3") {
        ret <- ret %>% 
          hc_colorAxis(min = -15e4, max = 15e4, tickInterval = 5e4)
      }
    }
    if (pct == "Yes") {
      ret <- ret %>% 
        hc_colorAxis(labels = list(format = paste0("{value}", number_unit())))
      if (level == "nuts1") {
        ret <- ret %>% 
          hc_colorAxis(min = -10, max = 10)
      }
      if (level == "nuts3") {
        ret <- ret %>% 
          hc_colorAxis(min = -40, max = 40, tickInterval = 20)
      }
    }
  }
  if (var == "In") {
    ret <- ret %>% 
      hc_colorAxis(stops = color_stops(
        colors = colorspace::divergingx_hcl(20, rev = TRUE)[11:20]
      ))
    if (pct == "No") {
      if (level == "nuts1") {
        ret <- ret %>% 
          hc_colorAxis(max = 4e5)
      }
      if (level == "nuts3") {
        ret <- ret %>% 
          hc_colorAxis(max = 3e5)
      }
    }
    if (pct == "Yes") {
      ret <- ret %>% 
        hc_colorAxis(labels = list(format = paste0("{value}", number_unit())))
      if (level == "nuts1") {
        ret <- ret %>% 
          hc_colorAxis(max = 15)
      }
      if (level == "nuts3") {
        ret <- ret %>% 
          hc_colorAxis(max = 50)
      }
    }
  }
  if (var == "Out") {
    ret <- ret %>% 
      hc_colorAxis(stops = color_stops(
        colors = colorspace::divergingx_hcl(20)[11:20]
      ))
    if (pct == "No") {
      if (level == "nuts1") {
        ret <- ret %>% 
          hc_colorAxis(max = 4e5)
      }
      if (level == "nuts3") {
        ret <- ret %>% 
          hc_colorAxis(max = 15e4)
      }
    }
    if (pct == "Yes") {
      ret <- ret %>% 
        hc_colorAxis(labels = list(format = paste0("{value}", number_unit())))
      if (level == "nuts1") {
        ret <- ret %>% 
          hc_colorAxis(max = 10)
      }
      if (level == "nuts3") {
        ret <- ret %>% 
          hc_colorAxis(max = 30)
      }
    }
  }
  ret
}
```

```{r laender-map-ui}
highchartOutput("laender_map")
```

```{r laender-map-srvr, context="server"}
output$laender_map <- renderHighchart({
  highchart() %>% 
    hc_add_series_map(
    map = nuts1_geojson,
    df = laender_df_trans() %>% 
      filter(year == input$year),
    value = input$plot_var,
    joinBy = "name",
    name = input$plot_var,
    dataLabels = list(enabled = TRUE, format = "{point.name}"),
    borderWidth = .3
  ) %>% 
    hc_map_colors(var = input$plot_var, pct = input$choose_pct, level = "nuts1") %>% 
    hc_legend(title = list(text = input$plot_var)) %>% 
    hc_tooltip(
      headerFormat = "{point.key}<br/>",
      pointFormat = paste0("<b>{point.value}", number_unit(), "</b><br/>")
    ) %>% 
    hc_add_theme(hc_theme(chart = list(style = list(
      fontFamily = "Fira Sans, sans-serif",
      fontVariantNumeric = "tabular-nums"
    )))) %>% 
    hc_exporting(
      enabled = TRUE,
      filename = "commuters_nuts1_map",
      formAttributes = list(target = "_blank"), # Prevents grey overlay after clicking
      buttons = list(contextButton = list(
        text = "Export",
        menuItems = c("downloadPNG", "downloadPDF", "downloadSVG")
      ))
    )
})
```

### `r emo::ji("chart_with_upwards_trend")` Time series

```{r laender-ts-ui}
highchartOutput("laender_ts")
```

```{r laender-ts-srvr, context="server"}
output$laender_ts <- renderHighchart({
  laender_df_trans() %>% 
    hchart("line", hcaes(x = year, y = !!sym(input$plot_var), group = name),
           marker = list(enabled = FALSE)) %>% 
    hc_xAxis(title = "") %>% 
    hc_tooltip(
      crosshairs = TRUE,
      headerFormat = "{point.key}<br/>",
      pointFormat = paste0("<b>{point.y}", number_unit(), "</b><br/>")
    ) %>% 
    hc_add_theme(hc_theme(chart = list(style = list(
      fontFamily = "Fira Sans, sans-serif",
      fontVariantNumeric = "tabular-nums"
    ))))
  # No exporting as still image. A non-interactive version of this chart would
  # not be very useful.
})
```

Column {data-width=3}
-----------------------------------------------------------------------

### Total {.value-box}

```{r calc-sum-fn, context="server"}
get_sum <- function(data, var) {
  if (input$choose_pct == "No") {
    value <- data %>% 
      filter(year == input$year) %>% 
      pull(.data[[var]]) %>% 
      sum() %>% 
      `/`(1e6) %>% 
      format(digits = 2, nsmall = 2) %>% 
      paste0(" M")
  } else {
    value <- data %>% 
      filter(year == input$year) %>% 
      summarise(sum = sum(.data[[var]]),
                pop = sum(pop)) %>% 
      mutate(pct = sum * 100 / pop) %>% 
      pull(pct) %>% 
      format(digits = 2, nsmall = 2) %>% 
      paste0("%")
  }
  paste0(var, ": ", value)
}
```

```{r total-box-nuts1-ui}
valueBoxOutput("nuts1_total_box_in")
valueBoxOutput("nuts1_total_box_out")
```

```{r total-box-nuts1-srvr, context="server"}
output$nuts1_total_box_in <- renderValueBox({
  valueBox(
    laender_df %>% 
      get_sum("In"),
    color = "#f8f8f8"
  )
})

output$nuts1_total_box_out <- renderValueBox({
  valueBox(
    laender_df %>% 
      get_sum("Out"),
    caption = paste0("Total in Germany ", input$year, "*"), color = "#f8f8f8",
    if (input$choose_pct == "No") {
      icon = "fa-user-friends"
    } else {
      icon = "fa-percentage"
    }
  )
})
```

### Description

<div class="description-text">Here, commuters are all employees subject to social insurance whose workplace is located in a different municipality from the one in which they live &ndash; regardless of whether and how often they actually commute. Looking at the number of commuters across state borders, there tends to have been more movement in and out of the states with a larger population, with the inflow generally exceeding the outflow in these states.

All Länder surrounding the three city states Berlin, Bremen and Hamburg have had more people commuting outwards than inwards in favour of the positive commuter balance in these city states. This pattern becomes even more striking when adjusted for the number of inhabitants.</div>

<span style="font-size:12px;line-height:normal">\* The employment statistics only account for commuters from abroad, but not those who commute to other countries for work, which results in a discrepancy between the total size of inflow and outflow.</span>

Between Kreise (NUTS-3) {data-icon="fa-list"}
=======================================================================

Column {data-width=7, .tabset}
-----------------------------------------------------------------------

### `r emo::ji("clipboard")` Table

```{r kreise-tbl-ui}
reactableOutput("kreise_tbl")
```

```{r kreise-tbl-srvr, context="server"}
output$kreise_tbl <- renderReactable({
  kreise_df_trans() %>% 
    filter(year == input$year) %>% 
    select(name, Out, Balance, In) %>% 
    reactable(
      defaultPageSize = 15, borderless = TRUE, wrap = FALSE, searchable = TRUE,
      defaultSorted = "Balance", defaultSortOrder = "desc",
      columns = list(
        name = colDef(name = "Kreis", defaultSortOrder = "asc",
                      maxWidth = 250),
        Out = colDef(
          cell = function(value) {
            width <- paste0(value * .75 * 100 / max(.$In), "%")
            value <- paste0(format(value, big.mark = " "), number_unit())
            bar_chart(value, width = width, fill = "#c7522b", align = "right")
          },
          align = "right"
        ),
        In = colDef(
          cell = function(value) {
            width <- paste0(value * .75 * 100 / max(.$In), "%")
            value <- paste0(format(value, big.mark = " "), number_unit())
            bar_chart(value, width = width, fill = "#008585", align = "left")
          },
          align = "left"
        ),
        Balance = colDef(
          name = emo::ji("balance_scale"),
          width = 80,
          cell = function(value) paste0(format(value, big.mark = " "), number_unit()),
          style = list(background = "#f8f8f8"),
          align = "right"
        )
      )
    )
})
```

### `r emo::ji("world_map")` Choropleth map

```{r kreise-map-ui}
highchartOutput("kreise_map")
```

```{r kreise-map-srvr, context="server"}
output$kreise_map <- renderHighchart({
  highchart() %>% 
    hc_add_series_map(
      map = nuts3_geojson,
      df = kreise_df_trans() %>% 
        filter(year == input$year),
      value = input$plot_var,
      joinBy = "hc-key",
      name = input$plot_var,
      dataLabels = list(enabled = TRUE, format = "{point.name}",
                        filter = list(operator = ">",
                                      property = "pop",
                                      value = "5e5")),
      borderWidth = .3
    ) %>% 
    hc_map_colors(var = input$plot_var, pct = input$choose_pct, level = "nuts3") %>% 
    hc_legend(title = list(text = input$plot_var)) %>% 
    hc_tooltip(
      headerFormat = "{point.key}<br/>",
      pointFormat = paste0("<b>{point.value}", number_unit(), "</b><br/>")
    ) %>% 
    hc_mapNavigation(enabled = TRUE) %>% 
    hc_add_theme(hc_theme(chart = list(style = list(
      fontFamily = "Fira Sans, sans-serif",
      fontVariantNumeric = "tabular-nums"
    )))) %>% 
    hc_exporting(
      enabled = TRUE,
      filename = "commuters_nuts1_map",
      formAttributes = list(target = "_blank"), # Prevents grey overlay after clicking
      buttons = list(contextButton = list(
        text = "Export",
        menuItems = c("downloadPNG", "downloadPDF", "downloadSVG")
      ))
    )
})
```

Column {data-width=3}
-----------------------------------------------------------------------

### Total {.value-box}

```{r total-box-nuts3-ui}
valueBoxOutput("nuts3_total_box_in")
valueBoxOutput("nuts3_total_box_out")
```

```{r total-box-nuts3-srvr, context="server"}
output$nuts3_total_box_in <- renderValueBox({
  valueBox(
    kreise_df %>% 
      get_sum("In"),
    color = "#f8f8f8"
  )
})

output$nuts3_total_box_out <- renderValueBox({
  valueBox(
    kreise_df %>% 
      get_sum("Out"),
    caption = paste0("Total in Germany ", input$year, "*"), color = "#f8f8f8",
    if (input$choose_pct == "No") {
      icon = "fa-user-friends"
    } else {
      icon = "fa-percentage"
    }
  )
})
```

### Description

<div class="description-text">Visualizations at the NUTS-3 level show how commuting centers around the larger cities/districts in each region. Looking at total inflow for 2019, the top five cities are Munich, Frankfurt a.M., Hamburg, Berlin and Cologne. They are also among those with the highest total outflow.

When scaling data in relation to population size, we can observe the important role of mid-sized cities for surrounding regions. This pattern becomes especially prominent when looking at choropleth maps for the balance of commuter flows (e.g. in Baden-Wuerttemberg and Bavaria). Also, with the statistics accounting for commuting between districts but still within state borders, Bremen and Hamburg are no longer outliers with regard to commuter inflow and balance, as was the case in the state-level comparison.</div>

<span style="font-size:12px;line-height:1.2">\* The employment statistics only account for commuters from abroad, but not those who commute to other countries for work, which results in a discrepancy between the total size of inflow and outflow.</span>
